---
title: "p8105_hw2_ml4198"
Author: Matt Lawlor
output: github_document
---

```{r}
library(tidyverse)
```

## Problem 1

Mr. Trashwheel datasheet

```{r}
mr_trashwheel = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) 
```

2017 Precipitation sheet

```{r}
precipitation_2017 = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

2018 Precipitation sheet

```{r}
precipitation_2018 = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

Recode months

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

Join precipitation sheets

```{r}
precipitation_data = 
  full_join(precipitation_2017, precipitation_2018)

precipitation_data =
  left_join(precipitation_data, month_df, by = "month")
```

The mr_trashwheel dataset includes 344 observations with key variables date, weight_tons, and volume_cubic_yards. The median number of sports balls in a dumpster in 2017 was `r mr_trashwheel %>% filter(year == 2017) %>%  pull(sports_balls) %>%  median()`.

The precipitation_data dataset includes 24 monthly observations of total monthly precipitation in years 2017 and 2018. The total precipitation in 2018 was `r precipitation_data %>%  filter(year == 2018) %>% pull(total) %>%  sum()`.

## Problem 2

NYC Transit subway dataset

```{r}
nyc_transit_df = read.csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>% 
    select(line:entry, vending, ada) %>% 
    mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) %>% 
    mutate_at(vars(route1:route11), replace_na, "None") %>% 
    mutate_at(vars(route8:route11), as.character()
    )
```

This data frame contains NYC transit lines, station names and locations, routes, entry types and whether they are ADA compliant. Data has been cleaned to include relevant variables as listed above and variable names have been standardized. There are 19 columns and 1868 rows in the data frame.

***Distinct lines***

```{r}
nyc_transit_df %>%   
  distinct(station_name, line)
```

There are 465 distinct stations across all lines.

***ADA compliant stations***

```{r}
nyc_transit_df %>% 
  pull(ada) %>% 
  sum()
```

There are 468 ADA compliant stations

***Proportion without vending allowing entrance***

```{r}
nyc_transit_df %>% 
  filter(vending == "NO") %>% 
  count(
    entry, 
    wt = NULL)
```

Of entrances/exits without vending, 37.7% allow entry.

***Pivot_longer routes***

```{r}
tidy_nyc_transit_df = 
  pivot_longer(
    nyc_transit_df,
    route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  )
```

***Distinct stations servng A train***

```{r}
tidy_nyc_transit_df %>%
  filter(route_name == "A") %>% 
  distinct(station_name)
```

56 stations serve the A train.