---
title: "p8105_hw2_ml4198"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
```

## Problem 1

Mr. Trashwheel datasheet

```{r}
mr_trashwheel = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) 
```

2017 Precipitation sheet

```{r}
precipitation_2017 = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

2018 Precipitation sheet

```{r}
precipitation_2018 = 
  readxl::read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

Recode months

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

Join precipitation sheets

```{r}
precipitation_data = 
  full_join(precipitation_2017, precipitation_2018)

precipitation_data =
  left_join(precipitation_data, month_df, by = "month")
```

The mr_trashwheel dataset includes 344 observations with key variables date, weight_tons, and volume_cubic_yards. The median number of sports balls in a dumpster in 2017 was `r mr_trashwheel %>% filter(year == 2017) %>%  pull(sports_balls) %>%  median()`.

The precipitation_data dataset includes 24 monthly observations of total monthly precipitation in years 2017 and 2018. The total precipitation in 2018 was `r precipitation_data %>%  filter(year == 2018) %>% pull(total) %>%  sum()`.




